I would like to express my deepest gratitude to Osamu Saotome, my thesis advisor at the Instituto Tecnológico de Aeronáutica, for his invaluable scientific guidance, continuous support, and the enriching discussions we shared throughout this research.

I am grateful to Roberto D’Amore for the VHDL lessons that inspired my interest in hardware description and the world of FPGAs during the EA-277 and EA-253 courses. My sincere thanks also go to Pedro Toledo for the fruitful partnership during the Master’s courses.

I would like to express my sincere gratitude to Johnny Öberg and Gracieth Batista from KTH Royal Institute of Technology for their valuable support and guidance during the course EA-251. Their insights and encouragement were essential in deepening my understanding of the subject and greatly contributed to the development of this work.

To my brother, Armando Zanone, I extend my appreciation for his support and assistance in revising the manuscript. I also remember my sister, Amanda, in memoriam, for all her encouragement to study and the help she provided during my engineering degree. I am deeply grateful to my parents, Armando and Orlanda, for their lifelong encouragement and unconditional love.

I would also like to thank my friends and colleagues at Omnisys Engenharia, especially Lucas Balabram, Marcos Dixini, and Marcelo Carlos, for the stimulating discussions on hardware description and FPGA design. Special thanks to Cassio Trolli for his friendship and encouragement.

Finally, I am profoundly grateful to my wife, Danila, and our three sons, Mateus, Rafael, and Daniel, for their patience, understanding, and unwavering support throughout the development of this thesis, especially during the times when my focus and presence were required elsewhere.